\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage[euler]{textgreek}
\usepackage{fancyhdr} % For header
\usepackage{mdframed} % for boxes
\usepackage{geometry} % For margins
\usepackage{amsthm} % For theorems
\usepackage{amsmath,amsfonts} % For math
\usepackage{mathtools}
 \geometry{
 a4paper,
 left=.75in,
 right=.75in,
 top=1in,
 }
 \newcommand{\bb}{\mathbb}
 \usepackage{sectsty}
\subsectionfont{\large\underline}
\fancyhf{}
\renewcommand\thesection{}
\renewcommand\thesubsection{}
\fancyhead[L]{\textbf{\leftmark\rightmark}}
\fancyhead[R]{\textbf{Summer writeup of research}}

\rfoot{Page \thepage}
\pagestyle{fancy}

\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\newtheorem*{example}{Example}
\newtheorem*{theorem}{Theorem}
\newtheorem*{lemma}{Lemma}
\newtheorem*{corollary}{Corollary}
\setlength{\parskip}{1em}
\setlength{\parindent}{0pt}
\begin{document}

\tableofcontents

\newpage

\section{Motivation}

Start writing what might be a good introduction for any paper or chapter.

Include citations!

\section{Definitions}

\begin{itemize}
  \item ASV
  \item OTU
  \item Zero-Inflation
  \item Phylogenetic tree
  \item Compositional
  \item Covariates/variables/ coefficients
\end{itemize}

\section{Notation}

\begin{itemize}
  \item Number of samples: $n$
  \item Sample index: $i: i = 1, \ldots, n$
  \item Number of ASVs: $k$
  \item ASV index: $j: j = 1, \ldots , k$
  \item Number of covariates: $q$. Note that $q$ does not include the intercept.
  \item Covariate index: $k: k = 1, \ldots , q$

  \item Responses: Transformed counts into relative abundances.

  Note that an alternative transformation to relative abundance might be good. %ADDCITE

   \begin{align*}
     y_{ij}\\
     \textbf{y}_{np \times 1} &= \begin{pmatrix} y_{i = 1, j = 1} \\\vdots \\ y_{i = 1, j = p}\\ \vdots \\ y_{i = n, j = 1} \\ \vdots  \\y_{i = n, j = p}\end{pmatrix}
   \end{align*}

   \item

   Design matrix $\mathbf{x}$ ('little x').

\begin{align*}
  \mathbf{x} &= \begin{pmatrix}x_{i=1, k = 1} & \cdots & x_{i = 1, k = q}\\
\vdots & \ddots & \\
x_{i = n, k = 1} & \cdots & x_{i = n, k = q}\end{pmatrix}
\end{align*}


    Since we don't have coefficient values measured on each ASV, they are only measured on each sample. Thus $x_{ij} = x_{i}$ for all $j = 1, \ldots , p$. Then we convert to design matrix big X:

   Design matrix $X$:

   \begin{align*}
     X &= x \otimes I_p\\
     &= \begin{pmatrix} x_{i = 1, k = 1} & 0 & \cdots & 0 \overbrace{\cdots}^p & 0  & x_{i = 1, k = q} & 0 & \cdots & 0 \\ 0 & x_{i = 1, k = 1} \\ x_{i = 2, k = 1} & 0 & \cdots & x_{i = 2, k = 2} & \cdots & x_{i = 2, k = q} & 0 & \cdots \end{pmatrix}
   \end{align*}


   \item Parameter vector $\beta$ with entries $\beta_{jk}$

   \begin{align*}
     \boldsymbol\beta_{pq \times 1} = \begin{pmatrix}
     \beta_{j = 1, k = 1}\\ \vdots \\
     \beta_{j = p, k = 1} \\ \vdots \\
     \beta_{j = 1, k = q} \\ \vdots \\
     \beta_{j = p, k = q}
   \end{pmatrix}
   \end{align*}


   \item Link:

   Link covariates to response.

   First assume that $\textbf{y}$ has the same mean and covariance structure as if $\textbf{y}_i \sim Dir(\alpha_{1p}, \ldots , \alpha_{ip})$ for all $i = 1, \ldots, n$.

   Then,

   \begin{align*}
     g(\alpha) = \log(\alpha) &= X \beta \\
     \log(\alpha_{ij}) &= x_{i} \beta_{ij}
   \end{align*}




\end{itemize}





\section{Dirichlet ideas }

Assume $\eta_i = (\eta_{i1}, \ldots , \eta_{ip}) \sim Dir(\alpha_{i1}, \ldots , \alpha_{ip})$, and $\sum_{j=1}^p \alpha_{ij} = \alpha_{i0}$

\subsection{Dirichlet Expected value}

\[E(\eta_{ij}) = \frac{\alpha_{ij}}{\alpha_{i0}} \]


\subsection{Dirichlet variance }

\begin{align*}
  Var(\eta_{ij}) &= \frac{\alpha_{ij}(\alpha_{i0} - \alpha_{ij})}{\alpha_{i0}^2(\alpha_{i0} + 1)}
\end{align*}

Note: $0 \leq Var(\eta_{ij}) \leq 1$

\subsection{Dirichlet covariance }

For $i \neq j$

\begin{align*}
  Cov(\eta_{is}, \eta_{it}) &=  - \frac{\alpha_{is}\alpha_{it}}{\alpha_{i0}^2(\alpha_{i0} + 1)}
\end{align*}


Note this matrix is singular.
Note this quantity is always negative

\subsection{Dirichlet correlation }

For $i \neq j$

\begin{align*}
  Cor(\eta_{is}, \eta_{it}) &= \frac{Cov(\eta_{is}, \eta_{it})}{\sqrt{Var(\eta_{is})Var(\eta_{it})}}\\
  &= -  \frac{\alpha_{is}\alpha_{it}}{\alpha_{i0}^2(\alpha_{i0} + 1)} \cdot \sqrt{\frac{\alpha_{i0}^2(\alpha_{i0} + 1)}{\alpha_{is}(\alpha_{i0} - \alpha_{is})}}\sqrt{\frac{\alpha_{i0}^2(\alpha_{i0} + 1)}{\alpha_{it}(\alpha_{i0} - \alpha_{it})}}\\
  &= - \sqrt{\frac{\alpha_{is}\alpha_{it}}{(\alpha_{i0} - \alpha_{is})(\alpha_{i0} - \alpha_{it})}}
\end{align*}


Note this is always negative

\section{GEE ideas}




(include definitions of all )

\section{Identifiability issues}

\end{document}
