\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage[euler]{textgreek}
\usepackage{fancyhdr} % For header
\usepackage{mdframed} % for boxes
\usepackage{geometry} % For margins
\usepackage{amsthm} % For theorems
\usepackage{amsmath,amsfonts} % For math
 \geometry{
 a4paper,
 left=.75in,
 right=.75in,
 top=1in,
 }
 \newcommand{\bb}{\mathbb}
 \usepackage{sectsty}
\subsectionfont{\large\underline}
\fancyhf{}
\renewcommand\thesection{}
\renewcommand\thesubsection{}
\fancyhead[L]{\textbf{\leftmark\rightmark}}
\fancyhead[R]{\textbf{Research TODO}}

\rfoot{Page \thepage}
\pagestyle{fancy}

% For checklist
\usepackage{enumitem,amssymb}
\newlist{todolist}{itemize}{2}
\setlist[todolist]{label=$\square$}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
\hspace{-2.5pt}}
\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}


\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\newtheorem*{example}{Example}
\newtheorem*{theorem}{Theorem}
\newtheorem*{lemma}{Lemma}
\newtheorem*{corollary}{Corrolary}
\setlength{\parskip}{1em}
\setlength{\parindent}{0pt}
\begin{document}

\section{Week 1}

\begin{itemize}
  \item Broad goal: Fix code.
  \begin{todolist}
    \item[\done] Use less filtered dataset
    \begin{itemize}
      \item Now at 30\% with 39 taxa.
    \end{itemize}
    \item[\done] Make graphs of parameter values at each iteration
    \begin{itemize}
      \item Did a lot of graphs, histograms and changes of betas between loops. Need to do final after changes are made.
    \end{itemize}
    \item[\done] Try stepwise update
    \begin{itemize}
      \item Using .1 as a constant. Still doesnt work, so likely a more complicated approach would not be helpful.
    \end{itemize}
    \item[\done] Fix problem where R inverse and others not updating in beta loop
    \item[\done] Change R-inv function to depend on alpha instead of X and beta since already calculated.
    \item[\done] Flip order of beta and phi loops.
    \item[\done] Look at residuals in each iteration. (in phi step)
    \item Keep something constant through loop and/or try using identity for R.
    \item[\done] Try different dataset? Maybe simulated one from previous paper.
    \begin{itemize}
      \item Same troubles
    \end{itemize}
    \item Distance matrix argument to dist?
    \begin{itemize}
      \item Need both forms, so may or may not be useful
    \end{itemize}
    \item Check if phi is correct now, and see how is used
    \item[\done] Figure out large alpha problem
    \begin{itemize}
      \item Seems the + and one loop beta fixed this?
    \end{itemize}
    \item Try a different covariate?
    \item Figure out why geem code uses + instead of - on update
    \begin{itemize}
      \item I think + is correct. At least it converges.
    \end{itemize}
    \item Identify which covariate is causing the trouble.
    \item Try removing the single OTU that seems to be contributing to problem.
    \item Save everything to list correctly to look at later.
    \begin{itemize}
      \item Saving rho, omega, phi, beta, diff (abs), num iterations
      \item Also do residuals.
    \end{itemize}
    \item Plot lists
    \item Remove scalar update.
    \item Write up algorithm for Yuan to go through? Ask to check.
    \item Final run through before meeting of graphs and numbers
  \end{todolist}
\end{itemize}



\end{document}
